Supplementary Material for `The response of trophic interaction networks to multiple stressors in a marine latitudinal gradient of the Southern Hemisphere'

### 2023-11-21 SupportingInformation.tex
New version to submit to GCB

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[vmargin=1in,hmargin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}

\usepackage{pdflscape,booktabs}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage[running]{lineno}
\linenumbers
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\hypertarget{the-response-of-trophic-interaction-networks-to-multiple-stressors-in-a-marine-latitudinal-gradient-of-the-southern-hemisphere}{%
\section{The response of trophic interaction networks to multiple
stressors in a marine latitudinal gradient of the Southern
Hemisphere}\label{the-response-of-trophic-interaction-networks-to-multiple-stressors-in-a-marine-latitudinal-gradient-of-the-southern-hemisphere}}

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Ecological networks are increasingly used as conceptual frameworks for
  understanding ecological community structure, identifying key species,
  and for ecosystem management and biodiversity conservation. Nowadays,
  there is a growing interest in understanding how ecological networks
  respond to environmental and anthropogenic stressors along
  geographical gradients.
\item
  We review for the first time the state-of-the-art knowledge on
  stressor response of marine food webs along the southwest Atlantic to
  Antarctic gradient (45 - 78ºS), considering the following study areas:
  San Jorge Gulf, Beagle Channel, Burdwood Bank, Scotia Sea, Potter Cove
  (Antarctica) and Weddell Sea (Antarctica). The aim of this review is
  threefold: 1) describe the complexity and structure of marine food
  webs along the southwest Atlantic to Antarctic axis from a network
  perspective; 2) identify the ongoing environmental and anthropogenic
  stressors for each marine ecosystem containing the food webs; and 3)
  resume proven food web changes and elaborate hypotheses on how the
  identified stressors might affect food web features, combining
  information on node- and network-level properties.
\item
  Environmental and anthropogenic stressor effects in the study gradient
  have been mostly assessed individually and at the organism and/or
  population, i.e., at the node level, with one exception: the effect of
  fisheries in the San Jorge Gulf food web. To address the most
  plausible stressors effects on the selected food webs, we built
  hypotheses for each study area considering the following: a)
  stressor(s), b) parameter(s) affected, c) node-level properties of the
  affected species, and d) network-level properties of the food web.
\item
  The most common stressor reported along the gradient is global
  warming, except for Beagle Channel and Burdwood Bank, which are more
  influenced by the introduction of an alien species and fisheries,
  respectively. We provide particular hypotheses on how warming might
  impact food webs.
\item
  We demonstrate the advantages of applying a network approach to better
  understand and predict the effect of stressors in marine ecosystems of
  the Southern Hemisphere.
\end{enumerate}

Keywords: anthropogenic stressors, environmental stressors, food webs,
latitudinal gradient, Southern Hemisphere

\hypertarget{introduction}{%
\subsection{1. Introduction}\label{introduction}}

The application of a network perspective has emerged as a powerful tool
to tackle the complexity of species interactions, facilitating a better
understanding of the structure and functioning of ecosystems (Belgrano,
Scharler, Scharler, Dunne, \& Ulanowicz, 2005; Thompson et al., 2012;
D'Alelio, Libralato, Wyatt, \& Ribera d'Alcalà, 2016). Trophic networks
(or food webs) allow identifying properties and key species that may be
crucial for ecosystem stability, and hence important for ecosystem
management and biodiversity conservation (Thompson et al., 2012). There
is a growing interest in understanding how ecological networks respond
to environmental and anthropogenic stressors along geographical
gradients (Baiser, Gotelli, Buckley, Miller, \& Ellison, 2012; Cirtwill,
Stouffer, \& Romanuk, 2015; Strydom et al., 2021; Bauer et al., 2022).
Yet, only a few studies have described variation in food web structure
along latitudinal gradients in marine ecosystems. The few that have come
from the Global North (Wood, Russell, Hanson, Williams, \& Dunne, 2015;
Kortsch et al., 2019; Frelat et al., 2022; Pecuchet et al., 2022),
whereas no studies, nor meta-analyses, on geographical variation in
marine food webs exist for the Global South (Southern Hemisphere).

Here we review for the first time the state-of-the-art knowledge on
stressor response of marine food webs along the southwest Atlantic to
Antarctic gradient (45 - 78ºS, Figure 1). We focus on proven and
expected changes in food webs driven by stressors in selected areas
along this large-scale latitudinal gradient. The aim of this review is
threefold: 1) describe the complexity and structure of marine food webs
along the southwest Atlantic to Antarctic axis from a network
perspective; 2) identify the ongoing environmental and anthropogenic
stressors for each marine ecosystem containing the food webs; and 3)
resume proven food web changes and elaborate hypotheses on how the
identified stressors might affect food web features (e.g.~energy flow,
stability), combining information on node- and network-level properties.
Finally, we suggest which additional data and analyses are needed to
gain insights into the stressors' effects on food web properties in the
southwest Atlantic to Antarctic region.

\hypertarget{the-structure-of-marine-food-webs-in-the-southwest-atlantic---antarctic-region}{%
\subsection{2. The structure of marine food webs in the Southwest
Atlantic - Antarctic
region}\label{the-structure-of-marine-food-webs-in-the-southwest-atlantic---antarctic-region}}

Together, the southwest and the Atlantic sector of the Southern Ocean
comprise one of the most biologically productive regions of the world's
oceans (Acha, Mianzan, Guerrero, Favero, \& Bava, 2004; Whitehouse et
al., 2008; Ballerini et al., 2014; Latorre et al., 2023). The referred
region extends from San Jorge Gulf (45ºS) in the Patagonian shelf to the
Weddell Sea (78ºS) in the Southern Ocean, and covers a well-connected
oceanic latitudinal gradient (R. P. Matano, Palma, \& Piola, 2010;
Steele, Thorpe, \& Turekian, 2010; Guihou, Piola, Palma, \& Chidichimo,
2020).

Throughout this latitudinal gradient, many investigations have been
carried out addressing the trophic ecology of specific species and
prey-predator relationships (Vinuesa \& Varisco, 2007; Sailley et al.,
2013; Pasotti et al., 2015; Riccialdelli, Newsome, Fogel, \& Fernández,
2017; Saunders, Hill, Tarling, \& Murphy, 2019; Riccialdelli et al.,
2020), but few studies have considered the complexity of the ecosystem
in terms of a high resolution of species and their prey-predator
interactions (but see Jacob et al. (2011), Marina et al. (2018), Funes,
Saravia, Cordone, Iribarne, \& Galván (2022), López-López, Genner,
Tarling, Saunders, \& O'Gorman (2022), Rodriguez, Marina, Schloss, \&
Saravia (2022), Marina et al. (in rev.)). Neglecting this complexity
might lead to a misunderstanding of the structure and functioning of the
ecosystems, and ultimately reduce the ability to predict ecosystem
responses to perturbations (Montoya, Woodward, Emmerson, \& Solé, 2009;
Woodward et al., 2010).

In this review, we consider marine areas in the Southern hemisphere for
which highly-resolved empirical food webs, in terms of species diversity
and trophic interactions, have been previously described. These areas
include: (1) San Jorge Gulf (45 - 47ºS, 65 - 68ºW), (2) Beagle Channel
(\textasciitilde54ºS, 68ºW), (3) Burdwood Bank (\textasciitilde54ºS,
59ºW), (4) Scotia Sea (48 - 58ºS, 50ºW), (5) Potter Cove (62ºS, 58ºW,
Antarctica), and (6) Weddell Sea (74 - 78ºS, 30ºW, Antarctica) (Figure
1). The geographic locations of these marine biomes span from temperate
to Antarctic, and are exposed to both environmental (sea warming,
glacial retreat) and anthropogenic (fishery, pollution) stressors.

\begin{figure}
\centering
\includegraphics[width=6.27in,height=3.528in]{Figures/Map_final.jpg}
\caption{Map of the study areas along the southwest Atlantic - Antarctic
latitudinal gradient. The areas are marked with numbers from one to six.
Smaller areas (1. San Jorge Gulf, 2. Beagle Channel, 3. Burdwood Bank,
and 5. Potter Cove) are shown on the panels to the left, whereas the
larger areas (4. North and South Scotia Sea, and 6. Weddell Sea) are
marked with a red rectangle on the map. The map was drawn using the
`marmap' R package (Pante, Simon-Bouhet, \& Irisson, 2023). Continental
contour shapefiles were obtained from www.ign.gob.ar.}
\end{figure}

San Jorge Gulf is the northernmost study area considered in this review.
It is a partially enclosed basin spanning approximately 34,000 km2 and
\textasciitilde100 m of maximum depths, located from Cabo dos Bahias
(44°55'S) to Cabo tres Puntas (47°06'S) (Figure 1). The shallower
northern and southern ends of the Gulf (with depths \textasciitilde40 m)
present two prominent frontal systems. These systems are the areas of
highest productivity (Bogazzi et al., 2005; Glembocki, Williams,
Góngora, Gagliardini, \& Orensanz, 2015). The Gulf's productivity
supports large invertebrate and vertebrate fisheries (María Eva Góngora,
González-Zevallos, Pettovello, \& Mendía, 2012), as well as marine
mammal and seabird populations (Yorio, 2009). The San Jorge Gulf food
web comprises 165 nodes and 1015 trophic interactions with a connectance
of 0.04. The percentage of top predators is 16\%, 78\% of intermediate
nodes and 6\% of basal nodes; 60\% of predators are omnivorous (Table
1). The most connected nodes are: the Argentine red shrimp Pleoticus
muelleri, the squat lobster Gimothea gregaria, squids (Illex argentinus
as dominant species) and Amphipoda. Notably, these nodes present
mid-trophic positions in the food web (3.0, 2.5, 3.6 and 2,
respectively) (Funes et al., 2022).

At a southern latitude, the Beagle Channel (54°52'S, 68°8'W) is an
interoceanic passage located at the southernmost tip of South America,
spanning 240 km in length and 5 km in width (\textasciitilde1200 km2),
and a range depth of 0 - 140 m (Figure 1). It features complex
coastlines, varying bathymetry, a prevailing west-to-east circulation
pattern, and a significant longitudinal gradient of glacial freshwater
discharge (Schloss et al., 2023). Bruno, Riccialdelli, Acha, \&
Fernández (2023) suggests that locally produced suspended particulate
organic matter (mostly composed by phytoplankton) and organic matter
accumulated in the sediments (macroalgae-originated detritus) are the
primary food sources for the marine ecosystem, as opposed to
allochthonous materials. Beagle Channel food web includes 145 nodes and
1115 trophic interactions with a connectance of 0.05 (Table 1). The food
web is suggested to have a wasp-waist structure, where the following
species play a crucial role in the dynamics of the ecosystem: Fueguian
sprat Sprattus fuegensis, longtail southern cod Patagonotothen ramsayi,
black southern cod P. tessellata, frogmouth Cottoperca trigloides, and
squat lobster Grimothea gregaria (Riccialdelli et al., 2020). Moreover,
the squat lobster has been identified as a species responsible for
linking modules and connecting the entire food web (Rodriguez et al.,
2022).

The so-called Burdwood Bank ecosystem comprises Marine Protected Areas
Namuncurá - Burdwood Bank I and II, meaning the shallow submarine
plateau named Burdwood Bank with a 200 m isobath boundary, and a deep
slope that reaches 4000 m in depth, respectively (Tombesi, Rabufetti, \&
Lovrich, 2020; Administración de Parques Nacionales, 2022) (Figure 1).
Physical features in the plateau are fairly stable, with salinity
averaging 34 all year round and temperature ranging between 4 and 8ºC
(Acha et al., 2004; Piola \& Falabella, 2009). The plateau is surrounded
by steep flanks of up to 4000 m depth, protected by the Marine Protected
Area Namuncurá - Burdwood Bank II (32,000 km2; \textasciitilde55º -
56ºS, \textasciitilde58º - 62ºW). Intense upwelling and mixing occur in
relation to the slope, entraining deep nutrient-rich waters into the
photic layer (Piola \& Falabella, 2009; Ricardo P. Matano, Palma, \&
Combes, 2019), and resulting in a fairly homogeneous water column both
spatially and temporally (Glorioso \& Flather, 1995; Ricardo P. Matano
et al., 2019). The Burdwood Bank food web comprises 379 nodes and 1788
interactions, with a connectance of 0.01, and an asymmetric degree
distribution (i.e.~most of the species have a relatively low number of
interactions and few species concentrate most of them). Almost half of
the consumers are omnivores (0.48), and the network displays a
small-world pattern (Marina et al., in rev.) (Table 1).

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[vmargin=1in,hmargin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{pdflscape,booktabs}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Supplementary Material for `The response of trophic interaction networks to multiple stressors in a marine latitudinal gradient of the Southern Hemisphere'},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Supplementary Material for `The response of trophic interaction
networks to multiple stressors in a marine latitudinal gradient of the
Southern Hemisphere'}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

Table S1. Node-level properties for species being impacted by
environmental and/or anthropogenic stressors (see Table 2 in the article
for details) in six study areas along the southwest Atlantic to
Antarctic gradient (45 - 78ºS). Degree: total number of interactions;
relative abundance: high (+++), medium (+), low (-), unknown (?).

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1809}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2340}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1383}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1277}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1383}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1596}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Food web}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Node / trophic species}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Degree}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Trophic level}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Omnivory level}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Relative abundance}
\end{minipage} \\
\midrule\noalign{}
\bottomrule\noalign{}
& \emph{Eleginops maclovinus} & 42 & 3.11 & 0.752 & ? \\
& \emph{Patagonotothen tessellata} & 35 & 3.29 & 0.721 & ++ \\
& \emph{Paralomis granulosa} & 26 & 3.28 & 0.277 & ? \\
& \emph{Patagonotothen cornucola} & 23 & 3.23 & 0.495 & +++ \\
& \emph{Nacella magellanica} & 22 & 2.46 & 0.397 & ? \\
& \emph{Mytilus edulis chilensis} & 21 & 2.42 & 0.368 & ? \\
& \emph{Sprattus fuegensis} & 17 & 3.05 & 0.422 & ++++ \\
& \emph{Harpagifer bispinis} & 15 & 3.51 & 0.145 & ? \\
& \emph{Paranotothenia magellanica} & 14 & 3.42 & 0.450 & +++ \\
& \emph{Odontesthes nigricans} & 11 & 3.50 & 0.428 & ? \\
& \emph{Patagonotothen sima} & 7 & 3.44 & 0.115 & ++ \\
& \emph{Grimothea gregaria} & 71 & 2.41 & 0.434 & ++ \\
& \emph{Sprattus fuegensis} & 42 & 3.45 & 0.688 & ++ \\
& \emph{Patagonotothen ramsayi} & 68 & 3.48 & 0.492 & +++ \\
& \emph{Cottoperca trigloides} & 20 & 4.16 & 0.742 & ++ \\
& \emph{Dissostichus eleginoides} & 49 & 4.01 & 0.797 & + \\
& \emph{Dorytheuthis gahi} & 38 & 3.45 & 0.896 & ? \\
& \emph{Henricia obesa} & 5 & 2.19 & 0.160 & ? \\
& \emph{Odontaster penicillatus} & 7 & 2.94 & 0.148 & ? \\
& \emph{Patagonotothen guntheri} & 63 & 3.39 & 0.390 & ++ \\
& \emph{Coelorinchus fasciatus} & 16 & 3.48 & 0.446 & ++ \\
& \emph{Coelorinchus marinii} & 8 & 3.32 & 0.658 & + \\
& \emph{Macrourus carinatus} & 17 & 4.13 & 0.455 & +++ \\
& \emph{Macrourus holotrachys} & 16 & 3.79 & 0.528 & +++ \\
& \emph{Tedania sp} & 2 & 2.00 & 0.000 & +++ \\
& \emph{Fusitriton magellanicus} & 8 & 3.11 & 0.017 & ++ \\
& \emph{Adelomelon ancilla} & 5 & 3.46 & 0.220 & + \\
& \emph{Zygochlamys patagonica} & 18 & 2.28 & 0.250 & ++ \\
& \emph{Hiatella sp} & 7 & 2.00 & 0.000 & ? \\
& \emph{Eurypodius latreillii} & 2 & 3.00 & 0.000 & + \\
& \emph{Sympagurus dimorphus} & 4 & 2.00 & 0.000 & ? \\
& \emph{Grimothea gregaria} & 32 & 2.50 & 0.250 & ? \\
& \emph{Libidoclaea granaria} & 7 & 2.00 & 0.000 & ? \\
& \emph{Alcyonium antarcticum} & 5 & 2.00 & 0.000 & ? \\
& \emph{Convexella magelhaenica} & 14 & 2.70 & 0.410 & ? \\
& \emph{Anthoptilum grandiflorum} & 9 & 2.78 & 0.395 & ? \\
& \emph{Primnoella scotiae} & 13 & 2.78 & 0.395 & ? \\
& \emph{Primnoisis antarctica} & 14 & 2.64 & 0.413 & ? \\
& \emph{Actinostola crassicornis} & 7 & 2.83 & 0.138 & + \\
& \emph{Brachiopoda spp} & 24 & 2.00 & 0.000 & ? \\
& \emph{Acodontaster sp} & 2 & 2.00 & 0.000 & ? \\
& \emph{Anasterias antarctica} & 2 & 3.57 & 0.000 & ++ \\
& \emph{Ceramaster patagonicus} & 4 & 3.53 & 0.096 & ? \\
& \emph{Henricia obesa} & 5 & 2.20 & 0.160 & ? \\
& \emph{Cosmasterias lurida} & 12 & 3.24 & 0.418 & ? \\
& \emph{Glabraster antarctica} & 4 & 2.00 & 0.000 & - \\
& \emph{Pteraster affinis} & 3 & 3.00 & 0.000 & ? \\
& \emph{Perknaster sp} & 2 & 3.08 & 0.000 & ? \\
& \emph{Ctenodiscus australis} & 2 & 2.00 & 0.000 & ? \\
& \emph{Odontaster sp} & 12 & 2.98 & 0.279 & ? \\
& \emph{Gorgonocephalus chilensis} & 4 & 3.59 & 0.009 & ? \\
& \emph{Ophiacantha vivipara} & 7 & 3.44 & 0.052 & ? \\
& \emph{Pseudechinus magellanicus} & 9 & 2.50 & 0.250 & ? \\
& \emph{Sterechinus agassizii} & 6 & 3.19 & 0.009 & ? \\
& \emph{Austrocidaris canaliculata} & 7 & 3.17 & 0.028 & - \\
& \emph{Chaetopterus antarcticus} & 1 & 2.00 & 0.000 & - \\
& \emph{Cnemidocarpa spp} & 12 & 2.10 & 0.090 & ? \\
& \emph{Daption capense} & 11 & 3.55 & 0.510 & ? \\
& \emph{Thalassarche melanophris} & 22 & 3.83 & 0.983 & ? \\
& \emph{Macronectes giganteus} & 7 & 3.93 & 0.903 & ? \\
& \emph{Thalassarche chrysostoma} & 16 & 4.06 & 0.906 & ? \\
& \emph{Diomedea epomophora} & 6 & 3.97 & 1.022 & ? \\
& \emph{Macronectes halli} & 6 & 4.00 & 1.017 & ? \\
& \emph{Procellaria aequinoctialis} & 8 & 4.06 & 0.782 & ? \\
& \emph{Euphausia superba} & 146 & 3.12 & 0.928 & +++ \\
& \emph{Dissostichus eleginoides} & 29 & 5.99 & 0.249 & +++ \\
& \emph{Krefftichthys anderssoni} & 157 & 4.50 & 1.010 & ++ \\
& \emph{Protomyctophum bolini} & 135 & 4.09 & 0.773 & + \\
& \emph{Electrona antarctica} & 192 & 4.50 & 1.000 & +++ \\
& \emph{Gymnoscopelus nicholsi} & 134 & 4.07 & 0.781 & - \\
& \emph{Gymnoscopelus braueri} & 185 & 4.53 & 0.978 & ++ \\
& \emph{Limacina retroversa} & 35 & 2.00 & 0.000 & +++ \\
& Benthic Diatomea & 36 & 1.00 & 0.000 & - \\
& Copepoda & 27 & 2.60 & 0.720 & - \\
& Phytoplankton & 23 & 1.00 & 0.000 & - \\
& Zooplankton & 17 & 2.80 & 0.640 & - \\
& Porifera & 18 & 2.00 & 8.043 & - \\
& \emph{Euphausia superba} & 17 & 3.42 & 0.734 & + \\
& Salpidae & 13 & 3.28 & 1.198 & - \\
& Ascidiacea & 9 & 2.45 & 0.607 & +++ \\
& \emph{Malacobelemnon daytoni} & 2 & 2.90 & 0.810 & + \\
Macroalgae & \emph{Palmaria decipiens} & 11 & 1.00 & 0.000 & + \\
& \emph{Gigartina skottsbergii} & 8 & 1.00 & 0.000 & + \\
& \emph{Desmarestia menziesii} & 7 & 1.00 & 0.000 & +++ \\
& \emph{Desmarestia anceps} & 6 & 1.00 & 0.000 & +++ \\
& \emph{Desmarestia antarctica} & 6 & 1.00 & 0.000 & + \\
& \emph{Iridaea cordata} & 6 & 1.00 & 0.000 & + \\
& \emph{Plocamium cartilagineum} & 6 & 1.00 & 0.000 & - \\
& \emph{Myriogramme manginii} & 4 & 1.00 & 0.000 & - \\
& \emph{Adenocystis utricularis} & 3 & 1.00 & 0.000 & + \\
& \emph{Georgiella confluens} & 3 & 1.00 & 0.000 & - \\
& \emph{Ascoseira mirabilis} & 2 & 1.00 & 0.000 & + \\
& \emph{Phaeurus antarcticus} & 2 & 1.00 & 0.000 & - \\
& \emph{Curdiea racovitzae} & 2 & 1.00 & 0.000 & - \\
& \emph{Callophyllis atrosanguinea} & 1 & 1.00 & 0.000 & - \\
& \emph{Neuroglossum delesseriae} & 1 & 1.00 & 0.000 & - \\
& \emph{Pantoneura plocamioides} & 1 & 1.00 & 0.000 & - \\
& \emph{Picconiella plumosa} & 1 & 1.00 & 0.000 & - \\
& \emph{Pleuragramma antarcticum} & 69 & 3.58 & 0.440 & +++ \\
& \emph{Chionodraco myersi} & 37 & 4.09 & 0.356 & ++ \\
& \emph{Pagetopsis maculatus} & 37 & 4.09 & 0.356 & + \\
& \emph{Gymnoscopelus braueri} & 62 & 3.53 & 0.274 & ? \\
& \emph{Dissostichus mawsoni} & 87 & 4.12 & 0.560 & ? \\
& Bacillariophyceae & 81 & 1.00 & 0.000 & +++ \\
& Silicoflagellates & 30 & 1.00 & 0.000 & ++ \\
& Foraminifera & 93 & 2.00 & 0.000 & ++ \\
& \emph{Euphausia superba} & 163 & 2.27 & 0.385 & +++ \\
& \emph{Salpa thompsoni} & 108 & 2.28 & 0.347 & ++ \\
& \emph{Leptonychotes weddellii} & 59 & 4.86 & 0.204 & ? \\
& \emph{Lobodon carcinophagus} & 28 & 4.24 & 0.553 & +++ \\
& \emph{Hydrurga leptonyx} & 67 & 4.72 & 0.346 & ? \\
& \emph{Arctocephalus gazella} & 61 & 4.67 & 0.331 & ? \\
& \emph{Thalassoica antarctica} & 19 & 4.32 & 0.429 & +++ \\
& \emph{Pagodroma nivea} & 11 & 4.21 & 0.209 & ++ \\
& \emph{Sterna paradisaea} & 7 & 4.04 & 0.347 & ++ \\
& \emph{Aptenodytes forsteri} & 53 & 4.78 & 0.211 & ++ \\
& \emph{Pygoscelis adeliae} & 7 & 3.78 & 0.282 & +++ \\
& \emph{Megaptera novaeangliae} & 4 & 3.26 & 0.014 & ? \\
& Hexactinellida & 49 & 2.00 & 0.000 & +++ \\
& The species of the San Jorge Gulf are not listed because the number of
stressed species reported is too high. At a functional group level of
resolution, the main impacted groups are: macroinvertebrates, fish, and
seabirds. & & & & \\
